---
title: "scot_covid"
author: "David Henderson"
date: "11/03/2020"
output: 
  html_document:
    theme: spacelab
    df_print: paged
    code_folding: hide
---


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(tweetrmd)

colours_davidhen <- c("#e53935", "#3949ab", "#8e24aa", "#039be5",
                      "#00897b", "#7cb342", "#fdd835", "#fb8c00",
                      "#6d4c41", "#546e7a")

theme_set(theme_minimal(base_family = "Roboto") +
            theme(panel.grid.minor = element_blank()))
```


# Plotting incidence of COVID-19 in Scotland

Incredibly, the best way top get this data is by manually updating your own spreadsheet from the daily release of figures. Many thanks to Josh Magennis who has been doing this....

```{r}
tweet_screenshot(tweet_url("GintoSmfc", "1238475613238898691"))
```

From that screenshot we can make a table to use for plotting

```{r}
scot_data <- tribble(~date, ~total_tests, ~tests_negative, ~deaths,
                     ymd(20200301), 698, 698, 0,
                     ymd(20200302), 816, 815, 0,
                     ymd(20200303), 915, 914, 0,
                     ymd(20200304), 1046, 1043, 0,
                     ymd(20200305), 1256, 1250, 0,
                     ymd(20200306), 1525, 1514, 0,
                     ymd(20200307), 1680, 1664, 0,
                     ymd(20200308), 1957, 1939, 0,
                     ymd(20200309), 2101, 2078, 0,
                     ymd(20200310), 2234, 2207, 0,
                     ymd(20200311), 2316, 2280, 0,
                     ymd(20200312), 2892, 2832, 0,
                     ymd(20200313), 3314, 3229, 0,
                     ymd(20200314), 3715, 3594, 0,
                     ymd(20200315), 4240, 4087, 1) %>% 
  mutate(new_test = total_tests - lag(total_tests), 
         tests_positive = total_tests - tests_negative, 
         daily_incidence = tests_positive - lag(tests_positive),
         pct_positive = round(tests_positive/total_tests*100, 3))

scot_data
```

And now plot the daily incidence


```{r, fig.width=9, fig.height=6.75, warning=FALSE}
scot_data %>% 
  ggplot(aes(date, daily_incidence)) +
  geom_col(fill = "#3949ab") +
  geom_point(aes(date, tests_positive)) +
  geom_path(aes(date, tests_positive), linetype = "dashed") +
  geom_text(aes(x = max(date) - days(1), y = max(tests_positive) - 5,
                label = "Line shows\ntotal cumulative\ncases"), 
            size = 2, fontface = "bold", hjust = 1) +
  geom_text(aes(x = ymd(20200312), y = 50, 
                label = "Bars show\ndaily new cases"),
            size = 2, fontface = "bold", hjust = 0) +
  scale_x_date(breaks = scales::pretty_breaks(n = 12),
               labels = scales::date_format(format = "%b-%d")) +
  scale_y_log10() +
  theme(plot.margin = margin(3,15,3,3,"mm")) +
  coord_cartesian(clip = "off") +
  labs(title = "Reported incidence of confirmed COVID-19 cases in Scotland",
       subtitle = "Logarithmic scale",
       y = "Number of cases\n(log scale)",
       x = "",
       caption = "Data derived from figures released by Scottish Government\n@_davidhen") -> plot_2_log
plot_2_log
  
```



```{r}
ggsave("plot_2_linear.png", plot_2_linear, width = 9, height = 6, dpi = 600)
ggsave("plot_2_log.png", plot_2_log, width = 9, height = 6, dpi = 600)
```


#UK cases

Data from https://www.arcgis.com/home/item.html?id=e5fd11150d274bebaaf8fe2a7a2bda11. Annoyingly, the "download" button doesn't have a seperate address so need to download manually and import that way. Also in .xlsx format. .csv would be preferable....


```{r}
uk_data <- readxl::read_excel("raw_data/uk_confirmed_cases_dept_h_sc.xlsx") %>% 
  mutate(date = ymd(DateVal)) %>% 
  rename(new_cases = CMODateCount,
         cum_cases = CumCases) %>% 
  select(date, everything(), -DateVal)
uk_data
```

```{r, fig.width=9, fig.height=6.75, warning=FALSE}
uk_data %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(fill = "#3949ab") +
  geom_point(aes(date, cum_cases)) +
  geom_path(aes(date, cum_cases), linetype = "dashed") +
  geom_text(aes(x = max(date) - days(3),
                y = max(cum_cases) - 10),
            label = "Cumulative total",
            size = 4, fontface = "bold") +
  scale_x_date(breaks = scales::pretty_breaks(n = length(uk_data$date)/2),
               labels = scales::date_format(format = "%b-%d"),
               guide = guide_axis(n.dodge = 2)) +
  scale_y_log10() +
  theme(plot.margin = margin(3,15,3,3,"mm")) +
  coord_cartesian(clip = "off") +
  labs(title = "Reported incidence of confirmed COVID-19 cases in UK",
       subtitle = "2020",
       y = "Number of cases\n(log scale)",
       x = "",
       caption = "Data from Dept of Health and Social Care\n@_davidhen") -> plot_3
plot_3
  
```


```{r}
ggsave("plot_3.png", plot_3, width = 12, height = 9, dpi = 600)
```


```{r}
regions <- read_csv("https://raw.githubusercontent.com/watty62/Scot_covid19/master/regional_cases.csv") %>% 
  pivot_longer(cols = 2:16, names_to = "hb", values_to = "new_cases") %>% 
  rename(date = Date) %>% 
  mutate(date = dmy(date)) %>% 
  group_by(hb) %>% 
  mutate(cum_cases = cumsum(new_cases)) %>% 
  ungroup %>% 
  arrange(hb, date)
regions
```


```{r, fig.width=9, fig.height=6}
regions %>% 
  filter(hb != "Grand Total") %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(fill = "#4477AA") +
  geom_point(aes(date, cum_cases)) +
  geom_path(aes(date, cum_cases)) +
  facet_wrap(~hb) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(3,15,3,3,"mm")) +
  labs(title = "COVID-19 confirmed cases by Scottish Health Board",
       subtitle = "Daily incidence (bars) and cumulative total (line)",
       x = "",
       y = "Number of confirmed cases",
       caption = "Data source: @watty62 https://github.com/watty62/Scot_covid19\nplot by @_davidhen https://github.com/davidhen/covid_19") -> hb_plot
hb_plot
```

```{r}
ggsave("hb_plot.png", hb_plot, width = 12, height = 9, dpi = 600)
```

